The Shape Matching Element Method for Meshless Animation of NURBs Models

### 2021-01-23 paper/sections/methods.tex
first pass done

\section{Methods}

@@

\subsection{Shape Matching}
\label{sec:shapematching}

@@

\subsection{Shape Matching Element Method}
 Each input object is composed of a set of boundary representations, $\Prts$, one for each part $\prt\in\Prts$. 
 Our kinematic mapping is a blended construction (as is done for skinning~\cite{10.1145/2614028.2615427}) which we write as

@@

 \subsubsection*{Multiple Part Projection}
As in ~\refsec{shapematching} we construct each $\Pi_i$ via shape matching. 
It is tempting to use an identical polynomial basis matrix, $\Pmatx$ for all parts, $\prt_i$, but recall that constructing $\Pmatx$ requires the selection of a deformation
origin $\bar{\refX}$~(\refeq{projection}). 
Using the same origin for all polynomials limits the expressivity of the kinematic model~\cite{STBS:2011}.
Instead we choose a number of points in $\refX$ ($|\Prts| > |\XBs| > 0$, where $\XBs$ is the set of points)  to act as deformation origins~(\refsec{origins}).
For each part we now construct $\Pmatxi$ using the associated deformation origin $\bar{\refX}_j\in \XBs$, where each origin can
be shared between multiple parts.

@@

\subsection{Mass Matrix}
The kinetic energy for a model may be expressed as
\begin{equation}
T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} d\Omega
\text{,}
\end{equation}
where $\Omega$ is the 3D integration domain, $\rho$ is the density, and $\dot{\mathbf{x}}$ is the velocity for some point in the domain. Using the deformation map defined in \ref{eqn:compact_x_map}, we may rewrite this kinetic energy as
\begin{equation}
\begin{split}
T & = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{q}}^T \mathbf{J}^T\mathbf{Y(X)}^T\mathbf{Y(X)}\ d\Omega \\
  & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{J}^T \left[ \int_{ \Omega} \rho \mathbf{Y(X)}^T\mathbf{Y(X)} d\Omega \right] \mathbf{J}\dot{\mathbf{q}} \\
  & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{M} \dot{\mathbf{q}}
\end{split}
\text{.}
\end{equation}
The second line is a result of the fact that all terms except for $\mathbf{Y(X)}$ are spatially constant, allowing us to move these outside the volume integral, which enables fast assembly of the mass matrix, $\mathbf{M}$.
\subsection{Error Energy}
Over the course of the simulation we may see the true boundary values deviate from their positions estimated by the polynomial. This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface. To address this we augment our kinetic energy with a term to account for this error. The error for a boundary point $\mathbf{x}_i^*$ that is known to lie on the limit surface of the NURBS can be written as
\begin{equation}
\begin{split}
\epsilon_i & = ||\mathbf{x}(\mathbf{X}_i) - \mathbf{x}_i^*|| \\
		   & = ||(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})\mathbf{\hat{b}}||
\end{split}
\text{,}
\end{equation}
where $\mathbf{S_i}$ extracts $\mathbf{x}_i^*$ from $\mathbf{\hat{b}}$. We modify this error further to be expressed in terms of the control points. We note that $\mathbf{x}^*(u,v)$ yields the position on the NURBS surface whereas $\mathbf{x(X)}$ is the deformed value estimated by the polynomial fitting. Furthermore $\mathbf{X}$ may be written as a function of the parametric coordinates $\mathbf{X}(u,v)$ as its position on the surface is produced using the initial set of control points representing the undeformed model. This leads to an energy potential over the NURBS surface in parametric space
\begin{equation}
T_E = \iint (\mathbf{x}(\mathbf{X}(u,v)) - \mathbf{x}^*(u,v))^T(\mathbf{x}(\mathbf{X}(u,v)) - \mathbf{x}^*(u,v)) du dv
\text{.}
\end{equation}
The discrete form of this written in terms of the control points may be written as
\begin{equation}
\begin{split}
T_E & = \sum_i^N  \dot{\mathbf{q}}^T\mathbf{J}^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})\mathbf{J}\dot{\mathbf{q}} \\
  & = \dot{\mathbf{q}}^T\mathbf{J}^T \left[ \sum_i^N  (\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i}) \right] \mathbf{J}\dot{\mathbf{q}} \\
  & = \dot{\mathbf{q}}^T \mathbf{M}_E \dot{\mathbf{q}}
\end{split}
\text{,}
\end{equation}
where $N$ is the total number of boundary points and $\mathbf{M}_E$ is our \textit{error mass matrix}.
% \subsection{Generalized Forces}
% We write our potential energy as 
% \begin{equation}
%     V = \int_\Omega \psi(\mathbf{F(X)}) d\Omega
% \text{.}
% \end{equation}
% To evaluate this energy (as well as its derivatives), we require a definition for the deformation gradient, $\mathbf{F(X)}$:
% \begin{equation}
%         \mathbf{F(X)} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
%                = \frac{\partial \mathbf{Y(X)}}{\partial \mathbf{X}}\mathbf{L}\hat{\mathbf{b}}
% \text{.}
% \end{equation}
% In the expression for the deformed position $\mathbf{x(X)}$ we see that the dependence on $\mathbf{X}$ only exists in $\mathbf{Y(X)}$, making $\mathbf{L}$ and $\hat{\mathbf{b}}$ constant terms when compute $\mathbf{F}$. Full details on how to compute $\mathbf{F}$ and its derivative $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ may be found in the appendix. (\myworries{I just wrote this because I'm too tired to right the d etails right now :p}. Now that we have an expression for the deformation gradient, we discretize our potential energy as follows
% \begin{equation}
%     V =  \sum_i^m \psi(\mathbf{F(X}_i) A_i
% \text{,}
% \end{equation}
% where $m$ is the number of integration points and $A_i$ is the volume for the $i$-th integration point.

% Like in the case of the kinetic energy, we also introduce an error force potential, which takes the simple form
% \begin{equation}
% V_E = \mathbf{q}^T \mathbf{M}_E \mathbf{q}
% \end{equation}
\subsection{Deformation Gradient}
At this point we have shown how to fit a polynomial describing the deformation of each surface. With these polynomial coefficients coupled with their associated centers of mass, we can construct an estimate of some arbitrary material point $\mathbf{X}$ using the deformation described by the surface. This lends itself to a general definition for the deformed position of a point as a weighted sum of the polynomials, which yields a unique polynomial for the point. Thus we arrive at a definition for the deformed position, $\mathbf{x}$, of a material point $\mathbf{X}$ in the undeformed space:
\begin{equation}
\label{eqn:deformation_map}
	\mathbf{x}(\mathbf{X}) = \sum_i^n w_i(\mathbf{X}) \left[ \mathbf{M}(\mathbf{X-\bar{X}}_j)\mathbf{c}_i + \mathbf{t}_j \right]
	\text{,}
\end{equation}
where we have $n$ surfaces, and $w_i$ and $\mathbf{c}_i$ are the weights and the coefficients for the $i$-th surface, respectively. The terms $\mathbf{\bar{X}}_j$ and $\mathbf{t}_j$ the undeformed and deformed centers of mass, respectively, that are associated with the $i$-th surface.

\subsection{External Forces}
\subsection{Time Integration}
With our definition of the deformation map as well as a strategy for integrating the volume, we now have the necessary ingredient to perform dynamic simulation. We follow the Variational Mechanics approach that defines a pair of kinetic and potential energies, and we apply the Euler-Lagrange equation to develop our equations of motion.

@@

%%% ------- OLD STUFF BELOW --------
% \section{Methods}
% \dave{We should probably put some pseudo code for the preprocessing and the runtime algorithms right up fron, along with a quick textual walk through of the method.}
% Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$. Following D-NURBS \cite{10.1145/176579.176580}, we take our degrees of freedom (DOF) to be control points of the surfaces and produce displacements directly on these control points at each simulation step. We note that we used a simplified version of D-NURBS in which the weights are held constant throughout the simulation and find this not to be limiting \myworries{(feel like we'll need better justification than saying just this?)}. We represent the degrees of freedom for the $i$-th NURBS surface with $m$ control points as $\mathbf{q}_i = \left[ \mathbf{p}_1^T, \dots, \mathbf{p}_m^T \right]^T \in \mathbb{R}^{3m}$. From this definition, we write the DOF for the entire model as $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ where $n$ is the number of NURBS surfaces for the model.

>>>>

%\section{Methods}
\section{Geometric Model}

@@

\section{Shape Matching}
\label{sec:shapematching}

@@

\section{Shape Matching Element Method}
 Each input object is composed of a set of boundary representations, $\Prts$, one for each part $\prt\in\Prts$. 
 Our kinematic mapping is a blended construction (as is done for skinning~\cite{10.1145/2614028.2615427}) which we write as

@@

 \subsection*{Multiple Part Projection}
As in ~\refsec{shapematching} we construct each $\Pi_i$ via shape matching. 
It is tempting to use an identical polynomial basis matrix, $\Pmatx$ for all parts, $\prt_i$, but recall that constructing $\Pmatx$ requires the selection of a deformation
origin $\bar{\refX}$~(\refeq{projection}). 
Using the same origin for all polynomials limits the expressivity of the kinematic model~\cite{STBS:2011}.
Instead we choose a number of points in $\refX$ ($|\Prts| > |\XBs| > 0$, where $\XBs$ is the set of points)  to act as deformation origins~(\refsec{origins}).
For each part we now construct $\Pmatxi$ using the associated deformation origin $\bar{\refX}_j\in \XBs$, where each origin can
be shared between multiple parts.

@@

\section{Physics-Based Animation}

\subsection{Mass Matrix}
The kinetic energy for a model may be expressed as
\begin{equation}
T = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{x}}^T\dot{\mathbf{x}} d\Omega
\text{,}
\end{equation}
where $\Omega$ is the 3D integration domain, $\rho$ is the density, and $\dot{\mathbf{x}}$ is the velocity for some point in the domain. Using the deformation map defined in \ref{eqn:compact_x_map}, we may rewrite this kinetic energy as
\begin{equation}
\begin{split}
T & = \frac{1}{2}\int_{ \Omega} \rho \dot{\mathbf{q}}^T \mathbf{J}^T\mathbf{Y(X)}^T\mathbf{Y(X)}\ d\Omega \\
  & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{J}^T \left[ \int_{ \Omega} \rho \mathbf{Y(X)}^T\mathbf{Y(X)} d\Omega \right] \mathbf{J}\dot{\mathbf{q}} \\
  & = \frac{1}{2} \dot{\mathbf{q}}^T \mathbf{M} \dot{\mathbf{q}}
\end{split}
\text{.}
\end{equation}
The second line is a result of the fact that all terms except for $\mathbf{Y(X)}$ are spatially constant, allowing us to move these outside the volume integral, which enables fast assembly of the mass matrix, $\mathbf{M}$.

\subsection{Error Energy}
\dave{We should cite VEM literature and point out this is a standard thing they do to make the algorithm consistent and convergent}
Over the course of the simulation we may see the true boundary values deviate from their positions estimated by the polynomial. This is most notable in cases where the degree of the polynomial on a NURBS is higher than the degree of the B-spline surface. To address this we augment our kinetic energy with a term to account for this error. The error for a boundary point $\mathbf{x}_i^*$ that is known to lie on the limit surface of the NURBS can be written as
\begin{equation}
\begin{split}
\epsilon_i & = ||\mathbf{x}(\mathbf{X}_i) - \mathbf{x}_i^*|| \\
		   & = ||(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})\mathbf{\hat{b}}||
\end{split}
\text{,}
\end{equation}
where $\mathbf{S_i}$ extracts $\mathbf{x}_i^*$ from $\mathbf{\hat{b}}$. We modify this error further to be expressed in terms of the control points. We note that $\mathbf{x}^*(u,v)$ yields the position on the NURBS surface whereas $\mathbf{x(X)}$ is the deformed value estimated by the polynomial fitting. Furthermore $\mathbf{X}$ may be written as a function of the parametric coordinates $\mathbf{X}(u,v)$ as its position on the surface is produced using the initial set of control points representing the undeformed model. This leads to an energy potential over the NURBS surface in parametric space
\begin{equation}
T_E = \iint (\mathbf{x}(\mathbf{X}(u,v)) - \mathbf{x}^*(u,v))^T(\mathbf{x}(\mathbf{X}(u,v)) - \mathbf{x}^*(u,v)) du dv
\text{.}
\end{equation}
The discrete form of this written in terms of the control points may be written as
\begin{equation}
\begin{split}
T_E & = \sum_i^N  \dot{\mathbf{q}}^T\mathbf{J}^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})\mathbf{J}\dot{\mathbf{q}} \\
  & = \dot{\mathbf{q}}^T\mathbf{J}^T \left[ \sum_i^N  (\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i})^T(\mathbf{Y}(\mathbf{X}_i)\mathbf{L} - \mathbf{S_i}) \right] \mathbf{J}\dot{\mathbf{q}} \\
  & = \dot{\mathbf{q}}^T \mathbf{M}_E \dot{\mathbf{q}}
\end{split}
\text{,}
\end{equation}
where $N$ is the total number of boundary points and $\mathbf{M}_E$ is our \textit{error mass matrix}.
% \subsection{Generalized Forces}
% We write our potential energy as 
% \begin{equation}
%     V = \int_\Omega \psi(\mathbf{F(X)}) d\Omega
% \text{.}
% \end{equation}
% To evaluate this energy (as well as its derivatives), we require a definition for the deformation gradient, $\mathbf{F(X)}$:
% \begin{equation}
%         \mathbf{F(X)} = \frac{\partial \mathbf{x}}{\partial \mathbf{X}}
%                = \frac{\partial \mathbf{Y(X)}}{\partial \mathbf{X}}\mathbf{L}\hat{\mathbf{b}}
% \text{.}
% \end{equation}
% In the expression for the deformed position $\mathbf{x(X)}$ we see that the dependence on $\mathbf{X}$ only exists in $\mathbf{Y(X)}$, making $\mathbf{L}$ and $\hat{\mathbf{b}}$ constant terms when compute $\mathbf{F}$. Full details on how to compute $\mathbf{F}$ and its derivative $\frac{\partial \mathbf{F}}{\partial \mathbf{q}}$ may be found in the appendix. (\myworries{I just wrote this because I'm too tired to right the d etails right now :p}. Now that we have an expression for the deformation gradient, we discretize our potential energy as follows
% \begin{equation}
%     V =  \sum_i^m \psi(\mathbf{F(X}_i) A_i
% \text{,}
% \end{equation}
% where $m$ is the number of integration points and $A_i$ is the volume for the $i$-th integration point.

% Like in the case of the kinetic energy, we also introduce an error force potential, which takes the simple form
% \begin{equation}
% V_E = \mathbf{q}^T \mathbf{M}_E \mathbf{q}
% \end{equation}
\subsection{Deformation Gradient}
\dave{gotta deal with this weight gradient}
At this point we have shown how to fit a polynomial describing the deformation of each surface. With these polynomial coefficients coupled with their associated centers of mass, we can construct an estimate of some arbitrary material point $\mathbf{X}$ using the deformation described by the surface. This lends itself to a general definition for the deformed position of a point as a weighted sum of the polynomials, which yields a unique polynomial for the point. Thus we arrive at a definition for the deformed position, $\mathbf{x}$, of a material point $\mathbf{X}$ in the undeformed space:
\begin{equation}
\label{eqn:deformation_map}
	\mathbf{x}(\mathbf{X}) = \sum_i^n w_i(\mathbf{X}) \left[ \mathbf{M}(\mathbf{X-\bar{X}}_j)\mathbf{c}_i + \mathbf{t}_j \right]
	\text{,}
\end{equation}
where we have $n$ surfaces, and $w_i$ and $\mathbf{c}_i$ are the weights and the coefficients for the $i$-th surface, respectively. The terms $\mathbf{\bar{X}}_j$ and $\mathbf{t}_j$ the undeformed and deformed centers of mass, respectively, that are associated with the $i$-th surface.

\subsection{External Forces}
\dave{Jacobian transpose}

\subsection{Time Integration}
With our definition of the deformation map as well as a strategy for integrating the volume, we now have the necessary ingredient to perform dynamic simulation. We follow the Variational Mechanics approach that defines a pair of kinetic and potential energies, and we apply the Euler-Lagrange equation to develop our equations of motion.

@@

\dave{Some extra notes on how we handle collisions, add a reference, just say we do this from here.}
%%% ------- OLD STUFF BELOW --------
% \section{Methods}
% \dave{We should probably put some pseudo code for the preprocessing and the runtime algorithms right up fron, along with a quick textual walk through of the method.}
% Given a volumetric model composed of one or more NURBS surfaces, the surface of each surface is reconstructed using a set of control points $\textbf{p}$ and their associated weights $w$. Following D-NURBS \cite{10.1145/176579.176580}, we take our degrees of freedom (DOF) to be control points of the surfaces and produce displacements directly on these control points at each simulation step. We note that we used a simplified version of D-NURBS in which the weights are held constant throughout the simulation and find this not to be limiting \myworries{(feel like we'll need better justification than saying just this?)}. We represent the degrees of freedom for the $i$-th NURBS surface with $m$ control points as $\mathbf{q}_i = \left[ \mathbf{p}_1^T, \dots, \mathbf{p}_m^T \right]^T \in \mathbb{R}^{3m}$. From this definition, we write the DOF for the entire model as $\mathbf{q} = \left[ \mathbf{q}_i^T, \dots, \mathbf{q}_n^T \right]^T$ where $n$ is the number of NURBS surfaces for the model.
